---
title: "R Notebook"
output:
  html_document: 
    keep_md: yes
  html_notebook: default
  pdf_document: default
---

```{r}
if (require("pacman")) {
    library(pacman)
  }else{
    install.packages("pacman")
    library(pacman)
  }
  pacman::p_load(tidyverse)
```

# Implementing Page Rank Methods

## Example Graph

Consider the following Graph taken from the paper:

```{r}
g1 <- igraph::graph.formula(1++2, 1+-8, 1+-5, 2+-5, 2+-7, 2+-8, 2+-6, 2+-9, 3++4, 3+-5, 3+-6, 3+-9, 3+-10, 4+-9, 4+-10, 4+-5, 5+-8, 6+-8, 7+-8)
plot(g1)
```

### Adjacency Matrix

The adjacency Matrix is given by:

```{r}
A <- igraph::get.adjacency(g1, names = TRUE, sparse = FALSE) %>% 
  as.matrix()

## Adjust the Order 
(A <- A[order(as.integer(row.names(A))), order(as.integer(colnames(A)))])
```

### State Distribution

The state distribution is the transpose of the adjacency matrix:

```{r}
(p0 <- t(A))
```

### Probability Transition Matrix

The probability transition matrix is such that each column of the initial state
distribution (i.e. the transposed adjacency matrix) is scaled to 1.

```{r}
p0 %*% diag(1/colSums(p0))
```
#### Create a Function

```{r}
adj_to_probTrans <- function(adjMat) {
  t(adjMat) %*% diag(1/colSums(t(adjMat)))
}

(T <- adj_to_probTrans(A)) %>% round(2)
```



## Page Rank Random Surfer

The random surfer page rank method modifies the probability transition matrix
$T$ so that the method works also for non-ergodic graphs by introducing
the possibility of a random jump, we'll call the surfer transition matrix $S$:

\begin{align}
    S &= \lambda T +  \left( 1- \lambda \right)B :\\
\ \\
    B&= \begin{bmatrix} 
	\frac{1}{N} & \frac{1}{N} & \ldots & \frac{1}{N} \\
	\frac{1}{N} & \frac{1}{N} & \ldots & \frac{1}{N} \\
        \vdots      & \vdots      & \ddots & \vdots \\
	\frac{1}{N} & \frac{1}{N} & \ldots & \frac{1}{N} \\
    \end{bmatrix}  \\
    N&= \left| \left| V \right| \right| \\
    \lambda &\in [0,1]
\end{align}


```{r}
B <- matrix(rep(1/nrow(T), length.out = nrow(T)**2), nrow = nrow(T))
l <- 0.8

S <- l*T+(1-l)*B
```


### Eigen Value Method

The eigenvector corresponding to the the eigenvalue of 1 will be the stationary point:

```{r}
eigen(S, symmetric = FALSE)
```

So in this case ~~the~~ a stationary point is $\langle -0.49, -0.53, -0.49, -0.48, -0.05, -0.05, -0.05, -0.04, -0.04, -0.04 \rangle$

which can be verified:

$$
1 \vec{p} = S\vec{p}
$$

```{r}
(p     <- eigen(S)$values[1] * eigen(S)$vectors[,1])
(p_new <- S %*% p)
```

However this vector does not sum to 1 so the scale should be adjusted (for probabilities the vector should sum to 1):

```{r}
(p_new <- p_new/sum(p_new))
```


### Power Value Method

Using the power method should give the same result, which it indeed does, but for the scale: 

```{r}
p_new <- p_new *123456789
    
while (sum(round(p, 9) != round(p_new, 9))) {
    (p     <- p_new) 
    (p_new <- S %*% p)
}

p_new
p

```

This answer is however identical in direction, if it scaled to 1 the same value will be returned:

```{r}
(p_new <- p_new/sum(p_new))
```


### Scaling

However if the initial state sums to 1, then the scale of the stationary vector will also sum to 1.

```{r}
p     <- c(1, 0, 0, 0, 0, 0, 0, 0, 0, 0)
p_new <- S %*% p
    
while (sum(round(p, 9) != round(p_new, 9))) {
    (p     <- p_new) 
    (p_new <- S %*% p)
}

cbind(p_new, p)
```






















































































