---
title: "Investigating Second Eigenvalue"
output: html_notebook
---

# Investigating Second Eigenvalue

```{r}
  if (require("PageRank")) {
      library(PageRank)
    }else{
      devtools::install_github("ryangreenup/PageRank")
      library(PageRank)
    }

  library(pacman)
  pacman::p_load(PageRank, devtools, Matrix, igraph, mise, tidyverse, rgl, latex2exp)
#  mise()
```



## Looking at Density

### Constants

Define some constants

```{r}
p    <- seq(from = 0.01, to = 0.99, length.out = 10)
beta <- seq(from = 1, to = 20, length.out = 4)
sz <- seq(from = 100, to = 10, length.out = 5)
input_var <- expand.grid("p" = p, "beta" = beta, "size" = sz)
input_var
```

### Function to Build Graph

```{r}
random_graph <- function(p, beta, size) {
      g1 <- igraph::erdos.renyi.game(n = size, p)
      A <- igraph::get.adjacency(g1) # Row to column
      A <- Matrix::t(A)

      A_dens <- mean(A)
      T      <- PageRank::power_walk_prob_trans(A, beta = beta)
      tr     <- sum(diag(T))
      e2     <- eigen(T, only.values = TRUE)$values[2] # R orders by descending magnitude
      return(c(abs(e2), mean(A), tr))
}
```

### Return results

Map the function

```{r}
nc <- length(random_graph(1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$p, X$beta, X$size)
  print(i/nrow(input_var))
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "A_dens", "trace")
(data <- cbind(input_var, Y)) %>% head()
```


### Plot Results


```{r}
pairs(data)
cor(data)
library(corrplot)
cormat = cor(data, method = 'spearman')
corrplot(cormat, method = "ellipse", type = "lower")
names(data)
```
Let's look at a 3d Output

```{r}
# plot3d(data$beta, data$A_dens, data$eigenvalue2, size = 2)
```



```{r}

names(data)
library(plotly)

d <- data
# d$beta <- log(d$beta)

fig <- plot_ly(d, x = ~A_dens, y = ~beta, z = ~eigenvalue2)
fig <- fig %>% add_markers(size = 1)
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Density'),
                     yaxis = list(title = 'Beta'),
                     zaxis = list(title = 'E2')))

fig

```

```{r}

names(data)
library(plotly)

#d <- data[sample(1:nrow(data), 1000),]
d <- data

fig <- plot_ly(d, x = ~A_dens, y = ~trace, z = ~eigenvalue2)
fig <- fig %>% add_markers(size = 1)
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Density'),
                     yaxis = list(title = 'trace'),
                     zaxis = list(title = 'E2')))

fig

```


Clearly I should be able to model this.

Let's look at density, it's continuous so I should be able to  take splices of A_dens.

A_dens is dependent on p, so I'll just use different ps


```{r}
p    <- seq(from = 0.1, to = 0.95, length.out = 5)
beta <- seq(from = 1, to = 10, length.out = 1000)
size = 100
input_var <- expand.grid("p" = p, "beta" = beta, "size" = size)
input_var

nc <- length(random_graph(1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$p, X$beta, X$size)
  print(i/nrow(input_var))
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "trace")
(data2 <- cbind(input_var, Y)) %>% head()
data2$p <- factor(data2$p)

```


```{r}
ggplot(data2[30:nrow(data2),], mapping = aes(col = p, x = beta, y = eigenvalue2)) +
  geom_point(size = 0.5) +
  stat_smooth() +
  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "Beta", y = TeX("Second Eigenvalue"), title = TeX("Second Eigenvalue given Matrix Density") ) +
  guides(col = guide_legend("Link Density"))  +
  theme_bw()
```

These look logarithmic, let's do a log transform:

```{r}

ggplot(data2[30:nrow(data2),], mapping = aes(col = p, x = log(beta), y = eigenvalue2)) +
  geom_point(size = 0.5) +
  stat_smooth() +
  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "Beta", y = TeX("Second Eigenvalue"), title = TeX("Second Eigenvalue given Matrix Density") ) +
  guides(col = guide_legend("Link Density"))  +
  theme_bw()
```
Link density is a function of size so E2\propto ln(Beta) is still pretty useful.

it seems that this is very nearly linear for low densities, the web would have a low density, for example ba graphs appear bounded above by 1/2:

```{r}
n_vec <- 0:20
m <- 3
y <- c()
for (n in n_vec) {
    x <- sample_pa(n = n, power = 3, m = m) %>% 
      get.adjacency() %>% 
      mean()
    y <- c(y, x)
}
plot(n_vec, y)
length(n_vec)
length(y)
```

So considering small Densities is wise, let's have a look.

So at this stage I would like to fit a model over what we have in the first plotly diagram, but, I'm going to look at ba graphs in case there is something more insightful.

```{r}

ggplot(data2[30:nrow(data2),], mapping = aes(col = factor(p), x = log(beta*p), y = eigenvalue2)) +
  geom_point(size = 0.5) +
  stat_smooth() +
  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "Beta", y = TeX("Second Eigenvalue"), title = TeX("Second Eigenvalue given Matrix Density") ) +
  
  guides(col = guide_legend("Link Density"))  +
  theme_bw()
```



It seems for very low densities that it is linear and then logarithmic for
moderate densities. BA models have density~=m/n so for the BA model we'll
just use beta.


```{r}

p       <- seq(from = 0.1, to = 0.95, length.out = 5)
size    <- seq(from = 100, to = 1000, length.out = 30)
beta <- seq(from = 1, to = 10, length.out = 1000)
beta <- 5
input_var <- expand.grid("p" = p, "beta" = beta, "size" = size)
input_var

nc <- length(random_graph(1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$p, X$beta, X$size)
  print(i/nrow(input_var))
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "trace")
(data2 <- cbind(input_var, Y)) %>% head()
data2$p <- factor(data2$p)



ggplot(data2, mapping = aes(col = factor(p), x = log(size), y = eigenvalue2)) +
  geom_point(size = 0.5) +
#  stat_smooth() +
#  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "size", y = TeX("Second Eigenvalue"), title = TeX("Second Eigenvalue given Matrix Density") ) +
  guides(col = guide_legend("Link Density"))  +
  theme_bw()
```




```{r}
p       <- seq(from = 0.1, to = 0.95, length.out = 40)
size    <- seq(from = 100, to = 1000, length.out = 5)
beta <- seq(from = 1, to = 10, length.out = 1000)
beta <- 5
input_var <- expand.grid("p" = p, "beta" = beta, "size" = size)
input_var

nc <- length(random_graph(1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$p, X$beta, X$size)
  print(i/nrow(input_var))
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "trace")
(data2 <- cbind(input_var, Y)) %>% head()
data2$p <- factor(data2$p)



ggplot(data2, mapping = aes(col = factor(p), x = log(size), y = eigenvalue2)) +
  geom_point(size = 0.5) +
#  stat_smooth() +
#  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "P", y = TeX("Second Eigenvalue"), title = TeX("Second Eigenvalue given Matrix Density") ) +
  guides(col = guide_legend("Size"))  +
  theme_bw()
```

