---
title: "Investigating Second Eigenvalue"
output: html_notebook
---

# Investigating Second Eigenvalue

```{r}
  if (require("PageRank")) {
      library(PageRank)
    }else{
      devtools::install_github("ryangreenup/PageRank")
      library(PageRank)
    }

  library(pacman)
  pacman::p_load(PageRank, devtools, Matrix, igraph, mise, tidyverse, rgl, latex2exp)
#  mise()
```



## Looking at Density

### Constants

Define some constants

```{r}
m    <- seq(from = 1, to = 10, length.out = 5)
m <- 5
beta <- seq(from = 1, to = 20, length.out = 10)
sz <- seq(from = 10, to = 1000, length.out = 10)
input_var <- expand.grid("m" = m, "beta" = beta, "size" = sz)
input_var
```

### Function to Build Graph

```{r}
random_graph <- function(m, beta, size) {
      g1 <- igraph::sample_pa(n = size, power = 3, m = m)
      A <- igraph::get.adjacency(g1) # Row to column
      A <- Matrix::t(A)

      A_dens <- mean(A)
      T      <- PageRank::power_walk_prob_trans(A, beta = beta)
      tr     <- sum(diag(T))
      e2     <- eigen(T, only.values = TRUE)$values[2] # R orders by descending magnitude
      return(c(abs(e2), mean(A), tr))
}
```

### Return results

Map the function

```{r}
rm(Y, data)
nc <- length(random_graph(1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$m, X$beta, X$size)
  print(i/nrow(input_var))
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "A_dens", "trace")
(data <- cbind(input_var, Y)) %>% head()
```


### Plot Results


```{r}
pairs(data)
cor(data)
library(corrplot)
cormat = cor(data, method = 'spearman')
corrplot(cormat, method = "ellipse", type = "lower")
names(data)
```
Let's look at a 3d Output

```{r}
names(data)
library(plotly)

# d <- data[sample(1:nrow(data), 100),]

fig <- plot_ly(d, x = ~A_dens, y = ~beta, z = ~eigenvalue2)
fig <- fig %>% add_markers(size = 1)
fig <- fig %>% layout(scene = list(xaxis = list(title = 'size'),
                     yaxis = list(title = 'Beta'),
                     zaxis = list(title = 'E2')))

fig

```

```{r}

names(data)
library(plotly)

# d <- data[sample(1:nrow(data), 1000),]

fig <- plot_ly(d, x = ~A_dens, y = ~trace, z = ~eigenvalue2)
fig <- fig %>% add_markers(size = 1)
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Density'),
                     yaxis = list(title = 'trace'),
                     zaxis = list(title = 'E2')))

fig

```


Clearly I should be able to model this.

Let's look at density, it's continuous so I should be able to  take splices of A_dens.

A_dens is dependent on p, so I'll just use different ps


```{r}
m    <- seq(from = 1, to = 9, length.out = 2)
beta <- seq(from = 1, to = 6, length.out = 10)
sz <- seq(from = 100, to = 1000, length.out = 4) %>% rev() # Big numbers first 

input_var <- expand.grid("m" = m, "beta" = beta, "size" = sz)
input_var


rm(Y, data)
nc <- length(random_graph(1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$m, X$beta, X$size)
  print(i/nrow(input_var))
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "A_dens", "trace")
(data2 <- cbind(input_var, Y)) %>% head()
# save(data2, file = "/home/ryan/Sync/Studies/2020Spring/DiscProj/data2_ba.rdata")
# load(file = "/home/ryan/Sync/Studies/2020Spring/DiscProj/data2_ba.rdata")
```


```{r}
data2$p <- (data2$m/data2$size)
names(data2)
ggplot(data2, mapping = aes(col = factor(p), x = beta, y = eigenvalue2)) +
  geom_point(size = 0.5) +
  stat_smooth(method = 'lm') +
  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "Beta", y = TeX("Second Eigenvalue"), title = TeX("Second Eigenvalue given Matrix Density") ) +
  guides(col = guide_legend("Link Density (by m/n)"))  +
  theme_bw()
```
It appears that m/n=density is its own feature, it does not affect E2 linearly so lets hold beta constant and see how it affects E2.

If for instance m/n increases, say quadratically for a constant beta,
then I should try and fit a linear model to (m/n)^2*beta

```{r}
m    <- seq(from = 1, to = 9, length.out = 20)
beta <- seq(from = 1, to = 6, length.out = 10)
beta <- 25
sz <- seq(from = 100, to = 1000, length.out = 10) %>% rev() # Big numbers first 

input_var <- expand.grid("m" = m, "beta" = beta, "size" = sz)
input_var


rm(Y, data)
nc <- length(random_graph(1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$m, X$beta, X$size)
  print(i/nrow(input_var))
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "A_dens", "trace")
(data2 <- cbind(input_var, Y)) %>% head()
```



```{r}
data2$p <- (data2$m/data2$size)
names(data2)
ggplot(data2, mapping = aes(col = factor(size), x = m/size, y = eigenvalue2)) +
  geom_point(size = 0.5) +
  stat_smooth(method = 'lm') +
  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "m/size", y = TeX("Second Eigenvalue"), title = TeX("Second Eigenvalue given Matrix Density") ) +
  guides(col = guide_legend("Size"))  +
  theme_bw()
```
So what about when I very m, i.e. lets map that to colour now:

```{r}
m <- 1:8
beta <- seq(from = 1, to = 6, length.out = 10)
beta <- 5
sz <- seq(from = 100, to = 1000, length.out = 20) %>% rev() # Big numbers first 

input_var <- expand.grid("m" = m, "beta" = beta, "size" = sz)
input_var


rm(Y, data)
nc <- length(random_graph(1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$m, X$beta, X$size)
  print(i/nrow(input_var))
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "A_dens", "trace")
(data2 <- cbind(input_var, Y)) %>% head()



data2$p <- (data2$m/data2$size)
names(data2)
ggplot(data2, mapping = aes(col = factor(m), x = m/size, y = eigenvalue2)) +
  geom_point(size = 0.5) +
  stat_smooth() +
  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "m/size", y = TeX("Second Eigenvalue"), title = TeX("Second Eigenvalue given Matrix Density") ) +
  guides(col = guide_legend("Vertices Added\n Each Step"))  +
  theme_bw()
```


Even though it's hard to capture the behaviour of m/n, the size of the graph and the average outdegree are going to be constants and so this relationship should persist
for graphs where m<<n.



actually if they are all linearly related, maybe I could use a multiple linear regression?

```{r}
m <- 1:8
beta <- seq(from = 1, to = 20, length.out = 100)
sz <- seq(from = 100, to = 1000, length.out = 20) %>% rev() # Big numbers first 

input_var <- expand.grid("m" = m, "beta" = beta, "size" = sz)
input_var


rm(Y, data)
nc <- length(random_graph(1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$m, X$beta, X$size)
  print(i/nrow(input_var))
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "A_dens", "trace")
(data2 <- cbind(input_var, Y)) %>% head()







mod <- lm(eigenvalue2 ~ m/size + size + m + beta, data2)
summary(mod)
plot(mod)
```





```{r}
mod <- lm(eigenvalue2 ~ m/size + m + size + beta , data = data2)
max(data2$m/data2$size)
summary(mod)
# layout(mat = matrix(1:4, nrow = 2))
# plot(mod)
# mean(mod$residuals**2)
hist(mod$residuals, breaks = 200)
```



Clearly non-normal residuals but +-5% is pretty good for a very complicated result.





I think I already looked at this but this again shows the relly clear, near linear relationship.


```{r}
ggplot(data2, mapping = aes(col = factor(m), x = beta, y = eigenvalue2)) +
  geom_smooth(size = 0.5) +
#  stat_smooth() +
  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "beta", y = TeX("Second Eigenvalue"), title = TeX("Second Eigenvalue given Matrix Density") ) +
  guides(col = guide_legend("Vertices Added\n Each Step"))  +
  theme_bw()
```

Obviously we can't know m, but, we can know the average out degree, so let's,
take a graph,
measure the average out degree,
use our model,
measure it's performance.



```{r}

```



























