---
title: "Eigenvalue vs density vs determinant"
author: "Ryan Greenup"
date: "10/4/2020"
output: html_document
---

# Investigating Second Eigenvalue

```{r}
  library(pacman)
  pacman::p_load(PageRank, devtools, Matrix, igraph, mise, tidyverse, rgl, latex2exp, plotly)
#  mise()
```

## Looking at Density

### Constants

Define some constants

```{r}
n <- 20
p <- 1:n/n
beta <- 1:n/n
beta <- runif(n)*100
#sz <- 1:n/n+10
sz <- (1:n/n)*100+10
input_var <- expand.grid("n" = n, "p" = p, "beta" = beta, "size" = sz)
head(input_var)
```

### Function to Build Graph

```{r}
random_graph <- function(n, p, beta, size) {
      g1 <- igraph::erdos.renyi.game(n = sz, p)
      A <- igraph::get.adjacency(g1) # Row to column
      A <- Matrix::t(A)

      A_dens <- mean(A)
      T      <- PageRank::power_walk_prob_trans(A)
      e2     <- eigen(T, only.values = TRUE)$values[2] # R orders by descending magnitude
      A_det  <- det(A)
      return(c(abs(e2), A_dens, A_det))
}
```

### Return results

Map the function

```{r}
nc <- length(random_graph(1, 1, 1, 1))
Y <- matrix(ncol = nc, nrow = nrow(input_var))
for (i in 1:nrow(input_var)) {
  X <- as.vector(input_var[i,])
  Y[i,] <-  random_graph(X$n, X$p, X$beta, X$size)
}
if (sum(abs(Y) != abs(Re(Y))) == 0) {
  Y <- Re(Y)
}
nrow(input_var)
nrow(Y)
Y <- as.data.frame(Y); colnames(Y) <- c("eigenvalue2", "density", "determinant")
(data <- cbind(input_var, Y)) %>% head()
data <- data[data$density!=0,]


```


### Plot Results


```{r}

chival <- dchisq(seq(from = 0, to = 40, length.out = 100), df = 10)*7
index  <- seq(from = 0, to = 2.2, length.out = 100)
chidata  <- data.frame(index = index, chi = chival)
ggplot(data) +
  geom_point(mapping = aes(x = density, y = eigenvalue2, size = beta, color = size, shape = factor(n))) +
  scale_size_continuous(range = c(0.1,1)) +
  labs(x = "Density of Adjacency Matrix", y = TeX("$\\xi_2$ of $T_{PW}$"))
```

This is interesting, lets look at a snapshot of it:

```{r}
pairs(data, cex = 0.01)
```

This suggests that only the determinant and density affect eigenvalue2 over all possible adjacency matrices.
the probability of forming edges is directly correlated with density so it can be ommitted, the plot of p~eigenvalue 2 is a heteroskedasdic inverse linear model, much like e2~density, suggesting that it woun't be of much predictive value compared to density. in practice we measure density not p so :shrug.

It appears that density ~ determinant is:

$$
\left|\mathrm{det}\left( \mathbf{A} \right)-0.5\right|\ <\ \left|\frac{1}{100\cdot \mathtt{mean}\left( \mathbf{A} \right)}\right|
$$

this restriction carries over to density, so the determinant can be used to indirectly predict e2.

extreme determinant implies that density will be about half, this means that a large determinant will mean a predictable albeit moderate E2. If det(A) = 0 we know nothing and must look at density. (is det faster or slower?)


Fairly pointless given that the determinant is harder to calculate anyway:

```{r}
library(Matrix)
n <- 1
sz <- 1000
p <- 0.4
dens_time <- replicate(n, {
  start <- Sys.time()
      g1 <- igraph::erdos.renyi.game(n = sz, p)
      A <- igraph::get.adjacency(g1) # Row to column
      mean_a <- mean(A)
  return(Sys.time()-start)
})

det_time<- replicate(n, {
  start <- Sys.time()
      g1 <- igraph::erdos.renyi.game(n = sz, p)
      A <- igraph::get.adjacency(g1) # Row to column
      det_a <- det(A)
  return(Sys.time() - start)
})
#mean((dens_time-det_time)/dens_time)
det_time_mean <- mean(det_time)
dens_time_mean <- mean(dens_time)

(dens_time_mean-det_time_mean)*100/dens_time_mean
```



So it seems that the only things that generally affect $\xi_2$ across all possible
adjacency matrices are density and determinant of $\mathbf{A}$, so let's look at
that:

```{r}
library(plotly)

mtcars$am[which(mtcars$am == 0)] <- 'Automatic'
mtcars$am[which(mtcars$am == 1)] <- 'Manual'
mtcars$am <- as.factor(mtcars$am)

d <- data[sample(1:nrow(data), 1000),]

fig <- plot_ly(d, x = ~determinant, y = ~density, z = ~eigenvalue2)
fig <- fig %>% add_markers(size = 1)
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Determinant'),
                     yaxis = list(title = 'density'),
                     zaxis = list(title = '2nd Eigenvalue')))

fig

```



